---
title: 'Fecal matter transplant in Huntington’s disease (Male)'
subtitle: "Data Analysis"
author: 'Saritha Kodikara'
#date: 'November 9, 2017'
date: '\today'
#date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document:
    toc: true   # can remove
    toc_depth: 3 # can remove
    number_sections: true
  highlight: zenburn
header-includes:
- \usepackage{fancyhdr}
- \usepackage{mathtools}
- \usepackage{xcolor, hyperref}
- \usepackage{lipsum}
- \usepackage{subfig}
- \usepackage{float}
- \setlength{\headheight}{34pt}
- \setlength{\footskip}{25pt}
- \pagestyle{fancy}
- \renewcommand{\headrulewidth}{0.5pt}
- \renewcommand{\footrulewidth}{0.5pt}
- \lhead{\includegraphics[width=8cm,height=1cm]{logo-unimelb}}
- \cfoot{\scriptsize School of Mathematics and Statistics, Melbourne Integrative Genomics \\ The University of Melbourne}
- \rhead{\thepage}
- \hypersetup{colorlinks   = true, linkcolor=blue, urlcolor  = blue}
- \fancypagestyle{plain}{\pagestyle{fancy}}
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
library(knitr)
# global options to show by default the code, dump the figures into /Figures etc
knitr::opts_chunk$set(dpi = 100, 
                      echo=TRUE, 
                      warning=FALSE, message=FALSE, eval = TRUE,
                      fig.show=TRUE, fig.width= 5,fig.height= 5,fig.align='center', out.width = '50%', fig.path= 'Figures/')
library(dplyr)
library(readxl)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(car)
library(nlme)
library(stargazer)
library(dotwhisker)
library(sjPlot)
require(xtable)
library(TSA)
library(jtools)
library(emmeans)
library(graphics)
library(tidyr)
library(ordinal)
library(viridis)
library(lmerTest)
library(ggpubr)
library(rstatix)
library(patchwork)
library(emmeans)
library(kableExtra)
library(MASS)
library(tibble)
```


<!-- the {-} indicate that you do not wish to number this section -->
# Executive Summary {-}

 - For repeated measures, Linear Mixed Models (LMM) were fitted with time as a covariate.
 - For one time point measures, Linear Mixed Models (LMM) were fitted with all two way interactions except for ATB*FMT (No observations for FMT and no ATB). However, only main effects were considered for variables that had few samples when covariates are considered together. For example IL17A-w14 had only 3 samples in HD-FMT group but when considered separately HD had 15 samples and FMT had 8 samples. Thus, IL17A-w14 was modeled without interactions.
 - Clasping score was modeled using a cumulative link mixed model.
 - For all models with significant effects on categorical variables (i.e, genotype, FMT, ATB), pairwise comparisons were performed using emmeans package.

# Data Input

```{r}
# load data from the csv file first. (the csv file is from your xlsx file)
data <- read_excel("Data/CarolGubert_FMT_male_data_July2021_final.xlsx")
data <- as.data.frame(data)
# extract group information
data$genotype <- factor(data$genotype,levels = c("WT", "HD"))
data$ATB<- factor(data$ATB,levels = c("NO", "YES"))
data$FMT<- factor(data$FMT,levels = c("NO", "YES"))
data$BOX <- factor(data$BOX)

Treatment<-ifelse((data$ATB=="YES"&data$FMT=="YES"),"ATB_FMT", 
                  ifelse((data$ATB=="YES"&data$FMT=="NO"),"Only ATB","No treatment"))%>% 
                factor(levels = c("No treatment","Only ATB","ATB_FMT"))
data.new<-add_column(data, Treatment = Treatment, .after = 5)
```

```{r,echo=FALSE, include=FALSE}
colnames(data.new)
```

# Linear Mixed Models - Repeated measures analysis

```{r}
data.repeated.1<-data.new[,c(1:6,8:37,61:75,91:116)]
long_data.repeated.1<-data.repeated.1 %>% 
  gather(v, value, BodyWeight_W6:FecalWaterContent_W20) %>% 
  separate(v, c("col", "Time"),sep="_W") %>% 
  arrange(ID) %>% 
  spread(col, value)
long_data.repeated.1$Time <-as.numeric(long_data.repeated.1$Time)
#filter out Time 6 and 7
long_data.repeated.1<-long_data.repeated.1%>%filter(Time!=6 &Time!=7)
long_data.repeated.1$Time<-long_data.repeated.1$Time-7

```

## LMM for Body Weight, Weight Gain, Rotarod, Fecal Output and Fecal water content

```{r}
#subset(long_data.repeated.1, !is.na(BodyWeight)) %>%
    #group_by(genotype, FMT, ATB) %>%
    #tally()


#subset(long_data.repeated.1, !is.na(BodyWeight)) %>%
    #group_by(genotype, Treatment, Time) %>%
    #tally()%>% summary()

#Interaction between ATB & FMT not considered as all no ATB had no FMT
BodyWeight.lmm <- lme(BodyWeight ~ genotype+Treatment+Time+
                         genotype:Treatment+genotype:Time+
                         Treatment:Time,random =(~1|BOX), 
                         data = subset(long_data.repeated.1, !is.na(BodyWeight)))

WeightGain.lmm <- lme(WeightGain ~ genotype+Treatment+Time+
                         genotype:Treatment+genotype:Time+
                         Treatment:Time,random =(~1|BOX), 
                         data = subset(long_data.repeated.1, !is.na(WeightGain)))

Rotarod.lmm <- lme(Rotarod ~ genotype+Treatment+Time+
                         genotype:Treatment+genotype:Time+
                         Treatment:Time,random =(~1|BOX),
                         data = subset(long_data.repeated.1, !is.na(Rotarod)))

FecalOutput.lmm <- lme(FecalOutput  ~ genotype+Treatment+Time+
                         genotype:Treatment+genotype:Time+
                         Treatment:Time,random =(~1|BOX),
                         data = subset(long_data.repeated.1, !is.na(FecalOutput)))

FecalWaterContent.lmm <- lme(FecalWaterContent  ~ genotype+Treatment+Time+
                         genotype:Treatment+genotype:Time+
                         Treatment:Time,random =(~1|BOX), 
                         data = subset(long_data.repeated.1, !is.na(FecalWaterContent)))

stargazer(BodyWeight.lmm ,WeightGain.lmm,Rotarod.lmm,
          FecalWaterContent.lmm,FecalOutput.lmm, type = "text",
          digits = 4,
          report = ('vc*p'),
          star.char = c("+", "*", "**", "***"),
          star.cutoffs = c(.1, .05, .01, .001),
          digit.separator = "", 
          notes = c("+ p<0.1; * p<0.05; ** p<0.01; *** p<0.001"), notes.append = FALSE)

```

## LMM for Food and Water intake  

```{r}
FoodIntake.data<-data.new[,c(1:6,38:49)]
long_FoodIntake.data<-FoodIntake.data %>% 
  gather(v, value, FoodIntake_W9:FoodIntake_W20) %>% 
  separate(v, c("col", "Time"),sep="_W") %>% 
  arrange(ID) %>% 
  spread(col, value)
long_FoodIntake.data$Time <-as.numeric(long_FoodIntake.data$Time)
long_FoodIntake.data$Time<-long_FoodIntake.data$Time-8

WaterIntake.data<-data.new[,c(1:6,50:60)]
long_WaterIntake.data<-WaterIntake.data %>% 
  gather(v, value, WaterIntake_W10:WaterIntake_W20) %>% 
  separate(v, c("col", "Time"),sep="_W") %>% 
  arrange(ID) %>% 
  spread(col, value)
long_WaterIntake.data$Time <-as.numeric(long_WaterIntake.data$Time)
long_WaterIntake.data$Time<-long_WaterIntake.data$Time-9
```

```{r}
FoodIntake.lmm <- lme(FoodIntake  ~ genotype+Treatment+Time+
                         genotype:Treatment+genotype:Time+
                         Treatment:Time,random =(~1|BOX),
                      data = subset(long_FoodIntake.data, !is.na(FoodIntake)))
WaterIntake.lmm <- lme(WaterIntake  ~ genotype+Treatment+Time+
                         genotype:Treatment+genotype:Time+
                         Treatment:Time,random =(~1|BOX),
                       data = subset(long_WaterIntake.data, !is.na(WaterIntake)))
stargazer(FoodIntake.lmm,WaterIntake.lmm, type = "text",
          digits = 4,
          report = ('vc*p'),
          star.char = c("+", "*", "**", "***"),
          star.cutoffs = c(.1, .05, .01, .001),
          digit.separator = "", 
          notes = c("+ p<0.1; * p<0.05; ** p<0.01; *** p<0.001"), notes.append = FALSE)
```

## LMM for CFC Conditioning and Extinction

```{r}
CFC.data<-data.new[,c(1:6,165:171)]
long_CFC.data<-CFC.data %>% 
  gather(v, value, CONDT_1:CONDT_6) %>% 
  separate(v, c("col", "Time"),sep="_") %>% 
  arrange(ID) %>% 
  spread(col, value)
long_CFC.data$Time <-as.numeric(long_CFC.data$Time)

CFC.ex.data<-data.new[,c(1:6,172:181)]
long_CFC.ex.data<-CFC.ex.data %>% 
  gather(v, value, `EXT_1-5`:`EXT_41-45`) %>% 
  separate(v, c("col", "Time"),sep="_") %>% 
  arrange(ID) %>% 
  spread(col, value)
long_CFC.ex.data$Time <- factor(long_CFC.ex.data$Time,ordered = TRUE,
                                     levels = c('1-5', '6-10', 
                                                '11-15', '16-20',
                                                '21-25', '26-30',
                                                '31-35', '36-40',
                                                '41-45'))
long_CFC.ex.data$Time = as.numeric(long_CFC.ex.data$Time)
```

```{r}
CFC_Conditioning.lmm <- lme(CONDT ~ genotype+Treatment+Time+CONDT_Base+
                         genotype:Treatment+genotype:Time+genotype:CONDT_Base+
                         Treatment:Time+Treatment:CONDT_Base+Time:CONDT_Base,
                         random=(~1|BOX),data =subset(long_CFC.data,!is.na(CONDT )))

CFC_Extinction.lmm <- lme(EXT ~ genotype+Treatment+Time+EXT_Base+
                         genotype:Treatment+genotype:Time+genotype:EXT_Base+
                         Treatment:Time+Treatment:EXT_Base+Time:EXT_Base,
                         random=(~1|BOX),data =subset(long_CFC.ex.data,!is.na(EXT )))

stargazer(CFC_Conditioning.lmm,CFC_Extinction.lmm, type = "text",
          digits = 4,
          report = ('vc*p'),
          star.char = c("+", "*", "**", "***"),
          star.cutoffs = c(.1, .05, .01, .001),
          digit.separator = "", 
          notes = c("+ p<0.1; * p<0.05; ** p<0.01; *** p<0.001"), 
          notes.append = FALSE)
```


# Linear Mixed Models - One Time point analysis

## Y-maze
```{r}
Ymaze.data<-data.new[,c(1:6,120:121)]
long_Ymaze.data<-Ymaze.data %>% 
  gather(v, value, Ymaze_novelArm:Ymaze_familiarArm) %>% 
  separate(v, c("col", "Arm"),sep="_") %>% 
  arrange(ID) %>% 
  spread(col, value)

long_Ymaze.data$Arm = factor(long_Ymaze.data$Arm,
                             levels = c("novelArm","familiarArm"),
                             labels = c("Novel","Familiar"))
```

```{r}
Ymaze.lmm <- lme(Ymaze ~ genotype+Treatment+Arm+
                         genotype:Treatment+genotype:Arm+
                         Treatment:Arm,random=list(BOX=~1, ID=~1),
                           data =subset(long_Ymaze.data,!is.na(Ymaze)))

stargazer(Ymaze.lmm, type = "text",
          digits = 4,
          report = ('vc*p'),
          star.char = c("+", "*", "**", "***"),
          star.cutoffs = c(.1, .05, .01, .001),
          digit.separator = "", 
          notes = c("+ p<0.1; * p<0.05; ** p<0.01; *** p<0.001"), 
          notes.append = FALSE)
```


## Phenotypic Variables (Gut Transit Time, Gut permeability, Cecum Weight, Cecum Length, ColonLength, Y-maze difference, Y-maze index, Y-maze Dist trial1, NOR, Brain Weight, Brake Ratio)

```{r}
dep_vars <-c("Ymaze_index","Ymaze_Dist_trial1","NOR_RI_5min",
             "BrainWeight","Digigait_PropelBrakeRatio_W14",
             "GutTransitTime_W14","GutTransitTime_W20","FITC_w14",
             "FITC_w20","CecumWeight_w14","CecumWeight_w20",
             "CecumLength_w14","CecumLength_w20",
             "ColonLength_w14","ColonLength_w20")

p_val<-lapply(dep_vars, function(r) {
  m <- lme(formula(paste(r, "genotype+Treatment+genotype:Treatment", sep = "~")),
           random=(~1|BOX),data = subset(data.new,!is.na(eval(parse(text = r)))))
  assign(paste0(r,".lmm_"),m, envir = .GlobalEnv)
  summary(m)$tTable[,5]})



stargazer(Ymaze_index.lmm_,Ymaze_Dist_trial1.lmm_,
          NOR_RI_5min.lmm_,BrainWeight.lmm_,Digigait_PropelBrakeRatio_W14.lmm_,
          type = "text",
          digits = 4,
          report = ('vc*p'),
          star.char = c("+", "*", "**", "***"),
          star.cutoffs = c(.1, .05, .01, .001),
          digit.separator = "", 
          notes = c("+ p<0.1; * p<0.05; ** p<0.01; *** p<0.001"), 
          notes.append = FALSE,
          dep.var.labels.include  = FALSE,
          column.labels=c("YmazeIndex","YmazeDist","NOR",
                          "BrainWeight","PropelBrakeRatio"),
          align=TRUE)

stargazer(GutTransitTime_W14.lmm_,
          GutTransitTime_W20.lmm_,FITC_w14.lmm_,FITC_w20.lmm_,
          type = "text",
          digits = 4,
          report = ('vc*p'),
          star.char = c("+", "*", "**", "***"),
          star.cutoffs = c(.1, .05, .01, .001),
          digit.separator = "", 
          notes = c("+ p<0.1; * p<0.05; ** p<0.01; *** p<0.001"), 
          notes.append = FALSE,
          dep.var.labels.include  = FALSE,
          column.labels=c("GutTransitTime-W14",
                          "GutTransitTime-W20","FITC-W14","FITC-W20"),
          align=TRUE)


stargazer(CecumWeight_w14.lmm_,CecumWeight_w20.lmm_,CecumLength_w14.lmm_,
          CecumLength_w20.lmm_,
          type = "text",
          digits = 4,
          report = ('vc*p'),
          star.char = c("+", "*", "**", "***"),
          star.cutoffs = c(.1, .05, .01, .001),
          digit.separator = "", 
          notes = c("+ p<0.1; * p<0.05; ** p<0.01; *** p<0.001"), 
          notes.append = FALSE,
          dep.var.labels.include  = FALSE,
          column.labels=c("CecumWeight-W14", "CecumWeight-W20",
                          "CecumLength-W14","CecumLength-W20"),
          align=TRUE)
stargazer(ColonLength_w14.lmm_,ColonLength_w20.lmm_,
          type = "text",
          digits = 4,
          report = ('vc*p'),
          star.char = c("+", "*", "**", "***"),
          star.cutoffs = c(.1, .05, .01, .001),
          digit.separator = "", 
          notes = c("+ p<0.1; * p<0.05; ** p<0.01; *** p<0.001"), 
          notes.append = FALSE,
          dep.var.labels.include  = FALSE,
          column.labels=c("ColonLength-W14","ColonLength-W20"),
          align=TRUE)
```


## Metabolic Variables (Acetate, Propionate, Isobutyrate,2-Methylbutyrate)

### LMM with only main effects as the number of samples were less than four for HD-FMT group

```{r}
#subset(data.new, !is.na(Acetate_w14)) %>%
    #group_by(genotype, Treatment) %>%
    #tally()
dep_vars <-c("Acetate_w14","Propionate_w14",
             "Isobutyrate_w14","Methylbutyrate_w14")

p_val<-lapply(dep_vars, function(r) {
  m <- lme(formula(paste(r, "genotype+Treatment", sep = "~")),
           random=(~1|BOX),data = subset(data.new,!is.na(eval(parse(text = r)))))
  assign(paste0(r,".lmm"),m, envir = .GlobalEnv)
  summary(m)$tTable[,5]})



stargazer(Acetate_w14.lmm,Propionate_w14.lmm,
          Isobutyrate_w14.lmm, Methylbutyrate_w14.lmm,
          type = "text",
          digits = 4,
          report = ('vc*p'),
          star.char = c("+", "*", "**", "***"),
          star.cutoffs = c(.1, .05, .01, .001),
          digit.separator = "", 
          notes = c("+ p<0.1; * p<0.05; ** p<0.01; *** p<0.001"), 
          notes.append = FALSE,
          dep.var.labels.include  = FALSE,
          column.labels=c("Acetate-14","Propionate-14",
                          "Isobutyrate-14","Methylbutyrate-14"),
          align=TRUE)

```



### With all two way interactions (except FMT*ATB)

```{r}
dep_vars <-c("Acetate_w20","Propionate_w20","Isobutyrate_w20",
             "Methylbutyrate_w20")

p_val<-lapply(dep_vars, function(r) {
  m <- lme(formula(paste(r, "genotype+Treatment+genotype:Treatment", sep = "~")),
           random=(~1|BOX),data = subset(data.new,!is.na(eval(parse(text = r)))))
  assign(paste0(r,".lmm"),m, envir = .GlobalEnv)
  summary(m)$tTable[,5]})


stargazer(Acetate_w20.lmm,Propionate_w20.lmm,
          Isobutyrate_w20.lmm,Methylbutyrate_w20.lmm,
          type = "text",
          digits = 4,
          report = ('vc*p'),
          star.char = c("+", "*", "**", "***"),
          star.cutoffs = c(.1, .05, .01, .001),
          digit.separator = "", 
          notes = c("+ p<0.1; * p<0.05; ** p<0.01; *** p<0.001"), 
          notes.append = FALSE,
          dep.var.labels.include  = FALSE,
          column.labels=c("Acetate-20","Propionate-20",
                          "Isobutyrate-20","Methylbutyrate-20"),
          align=TRUE)
```



## Cytokines Variables

### LMM with only main effects due to low samples

```{r}
#subset(data, !is.na(IFNg_w14)) %>%
  #  group_by(genotype, FMT, ATB) %>%
 #   tally()

dep_vars <-c("IL17A_w14", "IL17E_w14","IL21_w14","IL7R_w14" )

p_val<-lapply(dep_vars, function(r) {
  m <- lme(formula(paste(r, "genotype+Treatment", sep = "~")),
           random=(~1|BOX),data = subset(data.new,!is.na(eval(parse(text = r)))))
  assign(paste0(r,".lmm"),m, envir = .GlobalEnv)
  summary(m)$tTable[,5]})

stargazer(IL17A_w14.lmm, IL17E_w14.lmm,IL21_w14.lmm,IL7R_w14.lmm,
          type = "text",
          digits = 4,
          report = ('vc*p'),
          star.char = c("+", "*", "**", "***"),
          star.cutoffs = c(.1, .05, .01, .001),
          digit.separator = "", 
          notes = c("+ p<0.1; * p<0.05; ** p<0.01; *** p<0.001"), 
          notes.append = FALSE,
          dep.var.labels.include  = FALSE,
          column.labels=c( "IL17A-14", "IL17E-14","IL21-14","IL7R-14"),
          align=TRUE)

```



### With all two way interactions (except FMT*ATB)

```{r}
dep_vars <-c("IFNg_w14","IL1b_w14","IL22_w14","IL6_w14","TNFa_w14",
             "IFNg_w20","IL17A_w20","IL17E_w20","IL1b_w20",
             "IL21_w20","IL22_w20","IL6_w20","TNFa_w20")

p_val<-lapply(dep_vars, function(r) {
  m <- lme(formula(paste(r, "genotype+Treatment+genotype:Treatment", sep = "~")),
           random=(~1|BOX),data = subset(data.new,!is.na(eval(parse(text = r)))))
  assign(paste0(r,".lmm"),m, envir = .GlobalEnv)
  summary(m)$tTable[,5]})


stargazer(IFNg_w14.lmm,IL1b_w14.lmm,IL22_w14.lmm,IL6_w14.lmm,TNFa_w14.lmm,
          IFNg_w20.lmm,
          type = "text",
          digits = 4,
          report = ('vc*p'),
          star.char = c("+", "*", "**", "***"),
          star.cutoffs = c(.1, .05, .01, .001),
          digit.separator = "", 
          notes = c("+ p<0.1; * p<0.05; ** p<0.01; *** p<0.001"), 
          notes.append = FALSE,
          dep.var.labels.include  = FALSE,
          column.labels=c("IFNg-14","IL1b-14","IL22-14","IL6-14","TNFa-14",
                          "IFNg-20"),
          align=TRUE)

stargazer(IL17A_w20.lmm,IL17E_w20.lmm,IL1b_w20.lmm,IL21_w20.lmm,
          IL22_w20.lmm,
          type = "text",
          digits = 4,
          report = ('vc*p'),
          star.char = c("+", "*", "**", "***"),
          star.cutoffs = c(.1, .05, .01, .001),
          digit.separator = "", 
          notes = c("+ p<0.1; * p<0.05; ** p<0.01; *** p<0.001"), 
          notes.append = FALSE,
          dep.var.labels.include  = FALSE,
          column.labels=c("IL17A-20","IL17E-20","IL1b-20",
             "IL21-20","IL22-20"),
          align=TRUE)

stargazer(IL6_w20.lmm,TNFa_w20.lmm,
          type = "text",
          digits = 4,
          report = ('vc*p'),
          star.char = c("+", "*", "**", "***"),
          star.cutoffs = c(.1, .05, .01, .001),
          digit.separator = "", 
          notes = c("+ p<0.1; * p<0.05; ** p<0.01; *** p<0.001"), 
          notes.append = FALSE,
          dep.var.labels.include  = FALSE,
          column.labels=c("IL6-20","TNFa-20"),
          align=TRUE)

```




# Clasping score 

```{r}
data_clasping <- data.new[,c(1:6,78:90)] # extract the columns of data that will become Y
long_clasping<-data_clasping %>% 
  gather(v, value, Clasping_W8:Clasping_W20) %>% 
  separate(v, c("col", "Time"),sep="_W") %>% 
  arrange(ID) %>% 
  spread(col, value)
long_clasping$Time <-as.numeric(long_clasping$Time)-7
long_clasping$Clasping<-factor(long_clasping$Clasping, ordered = TRUE , levels = c(0:4))
t<-table(long_clasping$Time,long_clasping$Clasping)
t1<-table(long_clasping$Time,long_clasping$Clasping)%>%prop.table(margin = 1)
#subset(long_clasping, !is.na(Clasping)) %>%
#group_by(genotype, Treatment, Time) %>%
#tally() %>% summary()
```

```{r barplot_clas_count_male, echo = FALSE, fig.cap ='Bar plot for clasping over the 12 weeks', fig.pos="H",  fig.width = 10, fig.height = 9,  out.width = "1\\textwidth"}
ggplot(data.frame(as.matrix(t)), aes(fill=Var2, y=Freq, x=Var1)) + 
    geom_bar( stat="identity", position = position_stack(reverse = TRUE)) +
    scale_fill_viridis(discrete = T)+ylab("Count")+xlab("Time")+ 
  guides(fill=guide_legend(title="Clasping Score"))
```

```{r barplot_clas_per_male, echo = FALSE, fig.cap ='Bar plot for clasping over the 12 weeks', fig.pos="H",  fig.width = 10, fig.height = 9,  out.width = "1\\textwidth"}
ggplot(data.frame(as.matrix(t1)), aes(fill=Var2, y=Freq, x=Var1)) + 
    geom_bar( stat="identity", position = position_stack(reverse = TRUE)) +
    scale_fill_viridis(discrete = T)+ylab("Percent")+xlab("Time")+ 
  guides(fill=guide_legend(title="Clasping Score"))
```

```{r}
# HD mice clasping score
long_clasping_hd<-data_clasping %>% 
  gather(v, value, Clasping_W8:Clasping_W20) %>% 
  separate(v, c("col", "Time"),sep="_W") %>% 
  arrange(ID) %>% 
  spread(col, value)%>%filter(genotype=="HD")
long_clasping_hd$Time <-as.numeric(long_clasping_hd$Time)-7
long_clasping_hd$Clasping<-factor(long_clasping_hd$Clasping, ordered = TRUE , levels = c(0:4))
t1<-table(long_clasping_hd$Time,long_clasping_hd$Clasping)%>%prop.table(margin = 1)

```

```{r barplot_clas_per_male_hd, echo = FALSE, fig.cap ='Bar plot for clasping over the 12 weeks for HD mice', fig.pos="H",  fig.width = 10, fig.height = 9,  out.width = "1\\textwidth"}
ggplot(data.frame(as.matrix(t1)), aes(fill=Var2, y=Freq, x=Var1)) + 
    geom_bar( stat="identity", position = position_stack(reverse = TRUE)) +
    scale_fill_viridis(discrete = T)+ylab("Percent")+xlab("Time")+ 
  guides(fill=guide_legend(title="Clasping Score"))
```

```{r}
# WT mice clasping score
long_clasping_wt<-data_clasping %>% 
  gather(v, value, Clasping_W8:Clasping_W20) %>% 
  separate(v, c("col", "Time"),sep="_W") %>% 
  arrange(ID) %>% 
  spread(col, value)%>%filter(genotype=="WT")
long_clasping_wt$Time <-as.numeric(long_clasping_wt$Time)-7
long_clasping_wt$Clasping<-factor(long_clasping_wt$Clasping, ordered = TRUE , levels = c(0:4))
t1<-table(long_clasping_wt$Time,long_clasping_wt$Clasping)%>%prop.table(margin = 1)

```

```{r barplot_clas_per_male_wt, echo = FALSE, fig.cap ='Bar plot for clasping over the 12 weeks for WT mice', fig.pos="H",  fig.width = 10, fig.height = 9,  out.width = "1\\textwidth"}
ggplot(data.frame(as.matrix(t1)), aes(fill=Var2, y=Freq, x=Var1)) + 
    geom_bar( stat="identity", position = position_stack(reverse = TRUE)) +
    scale_fill_viridis(discrete = T)+ylab("Percent")+xlab("Time")+ 
  guides(fill=guide_legend(title="Clasping Score"))
```


## Ordinal regression

```{r}

Clasping.clm = clm(Clasping~genotype + Treatment + Time, 
           data=long_clasping, Hess=TRUE)
summary(Clasping.clm)

```

## Ordinal regression with random effect

```{r}

Clasping.clmm = clmm(Clasping~genotype + Treatment + Time+(1|BOX), 
           data=long_clasping, Hess=TRUE)
summary(Clasping.clmm)
```

## Comparison between Cumulative Link Model and Cumulative Link Mixed Model

```{r}
anova( Clasping.clmm,Clasping.clm)
```

Cumulative link mixed model is better than the cumulative link model with a lower AIC value.

# Post-Hoc test 

## Body Weight

```{r}
x<-emmeans(BodyWeight.lmm, pairwise ~ genotype*Treatment, 
           at = list(Time = c(1:13)))
df<-data.frame(x$contrasts)
that_cell<-df$p.value<0.05
kbl(x$contrasts, longtable=T, linesep = "", booktabs = T, digits = c(4,4,4,4,4,4),
    format = "latex",caption = "Post-hoc results for genotype *Treatment in body weight") %>%
  kable_styling(latex_options = c("repeat_header"))%>%
  column_spec(6,bold = that_cell,
              color =  ifelse(that_cell,"red","black"))%>%
  column_spec(1,color =  ifelse(that_cell,"red","black"))
```

## Weight Gain

```{r}
x<-emmeans(WeightGain.lmm, pairwise ~ genotype*Treatment, 
           at = list(Time = c(1:13)))
df<-data.frame(x$contrasts)
that_cell<-df$p.value<0.05
kbl(x$contrasts, longtable=T, linesep = "", booktabs = T, digits = c(4,4,4,4,4,4),
    format = "latex",caption = "Post-hoc results for genotype * Treatment in Weight Gain") %>%
  kable_styling(latex_options = c("repeat_header"))%>%
  column_spec(6,bold = that_cell,
              color =  ifelse(that_cell,"red","black"))%>%
  column_spec(1,color =  ifelse(that_cell,"red","black"))
```


## Fecal Water Content

```{r}
x<-emmeans(FecalWaterContent.lmm, pairwise ~ genotype*Treatment, 
           at = list(Time = c(1:13)))
df<-data.frame(x$contrasts)
that_cell<-df$p.value<0.05
kbl(x$contrasts, longtable=T, linesep = "", booktabs = T, digits = c(4,4,4,4,4,4),
    format = "latex",caption = "Post-hoc results for genotype * Treatment in Fecal Water Content") %>%
  kable_styling(latex_options = c("repeat_header"))%>%
  column_spec(6,bold = that_cell,
              color =  ifelse(that_cell,"red","black"))%>%
  column_spec(1,color =  ifelse(that_cell,"red","black"))
```


## Food Intake

```{r}
x<-emmeans(FoodIntake.lmm, pairwise ~ genotype*Treatment,
           at = list(Time = c(1:12)))
df<-data.frame(x$contrasts)
that_cell<-df$p.value<0.05
kbl(x$contrasts, longtable=T, linesep = "", booktabs = T, digits = c(4,4,4,4,4,4),
    format = "latex",caption = "Post-hoc results for genotype * Treatment in Food Intake") %>%
  kable_styling(latex_options = c("repeat_header"))%>%
  column_spec(6,bold = that_cell,
              color =  ifelse(that_cell,"red","black"))%>%
  column_spec(1,color =  ifelse(that_cell,"red","black"))
```

## Water Intake

```{r}
x<-emmeans(WaterIntake.lmm, pairwise ~ Treatment|genotype,
           at = list(Time = c(1:11)))
df<-data.frame(x$contrasts)
that_cell<-df$p.value<0.05
kbl(x$contrasts, longtable=T, linesep = "", booktabs = T, digits = c(4,4,4,4,4,4),
    format = "latex",caption = "Post-hoc results for Treatment given genotype in Water Intake") %>%
  kable_styling(latex_options = c("repeat_header"))%>%
  column_spec(7,bold = that_cell,
              color =  ifelse(that_cell,"red","black"))%>%
  column_spec(1,color =  ifelse(that_cell,"red","black"))
```


## CFC Extinction

```{r}
x<-emmeans(CFC_Extinction.lmm, pairwise ~ genotype|Treatment)
df<-data.frame(x$contrasts)
that_cell<-df$p.value<0.05
kbl(x$contrasts, longtable=T, linesep = "", booktabs = T, digits = c(4,4,4,4,4,4),
    format = "latex",caption = "Post-hoc results for genotype given Treatment in CFC Extinction") %>%
  kable_styling(latex_options = c("repeat_header"))%>%
  column_spec(7,bold = that_cell,
              color =  ifelse(that_cell,"red","black"))%>%
  column_spec(1,color =  ifelse(that_cell,"red","black"))
```

## Ymaze
```{r}
x<-emmeans(Ymaze.lmm, pairwise ~ genotype|Treatment*Arm)
df<-data.frame(x$contrasts)
that_cell<-df$p.value<0.05
kbl(x$contrasts, longtable=T, linesep = "", booktabs = T, digits = c(4,4,4,4,4,4),
    format = "latex",caption = "Post-hoc results for genotype given Treatment*Arm in Ymaze") %>%
  kable_styling(latex_options = c("repeat_header"))%>%
  column_spec(8,bold = that_cell,
              color =  ifelse(that_cell,"red","black"))%>%
  column_spec(1,color =  ifelse(that_cell,"red","black"))
```

## Brain Weight

```{r}
x<-emmeans(BrainWeight.lmm_, pairwise ~ genotype|Treatment)
df<-data.frame(x$contrasts)
that_cell<-df$p.value<0.05
kbl(x$contrasts, longtable=T, linesep = "", booktabs = T, digits = c(4,4,4,4,4,4),
    format = "latex",caption = "Post-hoc results for for genotype given Treatment in Brain weight") %>%
  kable_styling(latex_options = c("repeat_header"))%>%
  column_spec(7,bold = that_cell,
              color =  ifelse(that_cell,"red","black"))%>%
  column_spec(1,color =  ifelse(that_cell,"red","black"))
```

## Propel Brake Ratio- Week 14

```{r}
x<-emmeans(Digigait_PropelBrakeRatio_W14.lmm_, pairwise ~ genotype|Treatment)
df<-data.frame(x$contrasts)
that_cell<-df$p.value<0.05
kbl(x$contrasts, longtable=T, linesep = "", booktabs = T, digits = c(4,4,4,4,4,4),
    format = "latex",caption = "Post-hoc results for genotype given Treatment in Propel Brake Ratio") %>%
  kable_styling(latex_options = c("repeat_header"))%>%
  column_spec(7,bold = that_cell,
              color =  ifelse(that_cell,"red","black"))%>%
  column_spec(1,color =  ifelse(that_cell,"red","black"))
```

## FITC- Week 20

```{r}
x<-emmeans(FITC_w20.lmm_, pairwise ~ genotype|Treatment)
df<-data.frame(x$contrasts)
that_cell<-df$p.value<0.05
kbl(x$contrasts, longtable=T, linesep = "", booktabs = T, digits = c(4,4,4,4,4,4),
    format = "latex",caption = "Post-hoc results for genotype given Treatment in FITC") %>%
  kable_styling(latex_options = c("repeat_header"))%>%
  column_spec(7,bold = that_cell,
              color =  ifelse(that_cell,"red","black"))%>%
  column_spec(1,color =  ifelse(that_cell,"red","black"))
```

## CecumWeight- Week 14
```{r}
x<-emmeans(CecumWeight_w14.lmm_, pairwise ~ Treatment|genotype)
df<-data.frame(x$contrasts)
that_cell<-df$p.value<0.05
kbl(x$contrasts, longtable=T, linesep = "", booktabs = T, digits = c(4,4,4,4,4,4),
    format = "latex",caption = "Post-hoc results for Treatment given genotype in Cecum Weight") %>%
  kable_styling(latex_options = c("repeat_header"))%>%
  column_spec(7,bold = that_cell,
              color =  ifelse(that_cell,"red","black"))%>%
  column_spec(1,color =  ifelse(that_cell,"red","black"))
```




## Colon Length- Week 20

```{r}
x<-emmeans(ColonLength_w20.lmm_, pairwise ~ genotype|Treatment)
df<-data.frame(x$contrasts)
that_cell<-df$p.value<0.05
kbl(x$contrasts, longtable=T, linesep = "", booktabs = T, digits = c(4,4,4,4,4,4),
    format = "latex",caption = "Post-hoc results for genotype given Treatment in Colon length") %>%
  kable_styling(latex_options = c("repeat_header"))%>%
  column_spec(7,bold = that_cell,
              color =  ifelse(that_cell,"red","black"))%>%
  column_spec(1,color =  ifelse(that_cell,"red","black"))
```




## Acetate- Week 20
```{r}
x<-emmeans(Acetate_w20.lmm, pairwise ~ genotype*Treatment)
df<-data.frame(x$contrasts)
that_cell<-df$p.value<0.05
kbl(x$contrasts, longtable=T, linesep = "", booktabs = T, digits = c(4,4,4,4,4,4),
    format = "latex",caption = "Post-hoc results for genotype Treatment in Acetate") %>%
  kable_styling(latex_options = c("repeat_header"))%>%
  column_spec(6,bold = that_cell,
              color =  ifelse(that_cell,"red","black"))%>%
  column_spec(1,color =  ifelse(that_cell,"red","black"))
```

## IL17E- Week 20
```{r}
x<-emmeans(IL17E_w20.lmm, pairwise ~ Treatment|genotype)
df<-data.frame(x$contrasts)
that_cell<-df$p.value<0.05
kbl(x$contrasts, longtable=T, linesep = "", booktabs = T, digits = c(4,4,4,4,4,4),
    format = "latex",caption = "Post-hoc results for Treatment given genotype in IL17E") %>%
  kable_styling(latex_options = c("repeat_header"))%>%
  column_spec(7,bold = that_cell,
              color =  ifelse(that_cell,"red","black"))%>%
  column_spec(1,color =  ifelse(that_cell,"red","black"))
```

# Information about the used R session

```{r}
sessionInfo()
```